```{r setup}
library(dplyr)
```

## Filtering

Filtering student performance based on lunch status

```{r}
filter(studentdata, lunch == "standard")
filter(studentdata, lunch == "free/reduced")
```

Filtering student performance based on parental education level

```{r}
parent <- studentdata$parental.level.of.education
filter(studentdata, parent == "some high school")
filter(studentdata, parent == "some college")
filter(studentdata, parent == "associate's degree")
filter(studentdata, parent == "bachelor's degree")
filter(studentdata, parent == "master's degree")
```

## Selecting (Creating smaller data sets)

Data set only including parental education level and student performance metrics

```{r}
select(studentdata, parental.level.of.education, math.score, reading.score, writing.score)
```

Data set only including race/ethnicity and student performance metrics

```{r}
select(studentdata, race.ethnicity, math.score, reading.score, writing.score)
```

Data set only including test prep status and student performance metrics

```{r}
select(studentdata, test.preparation.course, math.score, reading.score, writing.score)
```

## Mutating

Data set including student performance metrics and their means

```{r}
metrics <- select(studentdata, math.score, reading.score, writing.score)

mutate(metrics, 
       math_mean = mean(math.score, na.rm=T),
       reading_mean = mean(reading.score, na.rm=T),
       writing_mean = mean(writing.score, na.rm=T))
```

## Summaries

Summarizing metrics by parental level of education

```{r}
summarize(studentdata, 
          mean_math = mean(math.score, na.rm=T),
          mean_reading = mean(reading.score, na.rm=T),
          mean_writing = mean(writing.score, na.rm=T),
          .by = parental.level.of.education)
```

## Arranging

Arranging math scores from lowest to highest

```{r}
arrange(studentdata, math.score)
```

Arranging math scores from highest to lowest

```{r}
arrange(studentdata, desc(math.score))
```

## Creating a new visualization

Scatter Plot describing math scores with respect to test prep course status

```{r}
library(ggplot2)

ggplot(data=studentdata, aes(x=test.preparation.course, y=math.score))+
       geom_jitter(aes(color = test.preparation.course), width=0.25)+
  labs(title = "Math Scores with Respect to Test Preparation Course Status",
       x="Test Preparation Course Status",
       y= "Math Score")
```
