## Lab 7: Hypothesis Testing 

#### Load required packages

```{r setup}
library(dplyr)
library(tidyr)
library(janitor)
```

#### Read in data set

```{r}
read.csv("DSRP2023/Data/StudentsPerformance.csv")

studentdata <- read.csv("DSRP2023/Data/StudentsPerformance.csv")
```

#### Establish variables and hypotheses

Categorical Variables: Parental level of education (master's degree, some high school)

Numerical Variable: math score

**Null hypothesis**: The average math score of students whose parents have a master's degree is equal to the average math score of students whose parents have only completed up to some high school.

**Alternate hypothesis**: The average math score of students whose parents have a master's degree is different from the average math score of students whose parents have only completed up to some high school.

This test will be an unpaired, two-sided T-test.

#### Create new, sub-datasets

```{r}
MastersMath <- studentdata |> filter (parental.level.of.education == "master's degree")|>
  select(parental.level.of.education, math.score)

SomeHSMath <- studentdata |> filter (parental.level.of.education == "some high school") |>
  select (parental.level.of.education, math.score)

```

#### Run hypothesis test

```{r}
t.test(MastersMath$math.score, SomeHSMath$math.score, paired = F, alternative = "two.sided")
```

The p-value is 0.007839, which is lower than the established level of significance (0.05). There is significant statistical evidence to suggest that there is a difference between the mean math score of students whose parents have a master's degree and the mean math score of students whose parents have only completed up to some high school.
