---
---
---

# Lab 7: Hypothesis Testing

## Part One

#### Load required packages

```{r setup}
library(dplyr)
library(tidyr)
library(janitor)
library(corrplot)
```

#### Read in data set

```{r}
studentdata <- read.csv("../Data/StudentsPerformance.csv")
```

#### Establish variables and hypotheses

Categorical Variables: Parental level of education (master's degree, some high school)

Numerical Variable: math score

**Null hypothesis**: The average math score of students whose parents have a master's degree is equal to the average math score of students whose parents have only completed up to some high school.

**Alternate hypothesis**: The average math score of students whose parents have a master's degree is different from the average math score of students whose parents have only completed up to some high school.

This test will be an unpaired, two-sided T-test.

#### Create new, sub-datasets

```{r}
MastersMath <- studentdata |> filter (parental.level.of.education == "master's degree")|>
  select(parental.level.of.education, math.score)

SomeHSMath <- studentdata |> filter (parental.level.of.education == "some high school") |>
  select (parental.level.of.education, math.score)

```

#### Run hypothesis test

```{r}
t.test(MastersMath$math.score, SomeHSMath$math.score, paired = F, alternative = "two.sided")
```

The p-value is 0.007839, which is lower than the established level of significance (0.05). There is significant statistical evidence to suggest that there is a difference between the mean math score of students whose parents have a master's degree and the mean math score of students whose parents have only completed up to some high school.

## Part Two

#### Establish Variables and Hypotheses

Categorical Variable: Parental Level of Education (some high school, some college, associate's degree, bachelor's degree, master's degree)

Numeric: Writing score

**Null hypothesis**: The average writing score of students is equal across all levels of parental education.

**Alternate hypothesis**: The average writing score of students differs between levels of parental education.

#### Conduct ANOVA Test

```{r}
anova_result <- aov(data = studentdata, writing.score ~ parental.level.of.education)

summary(anova_result)

TukeyHSD(anova_result)
```

The p-value of the ANOVA test is 1.12\*10\^-13, which is significantly lower than the level of significance (0.05). There is sufficient evidence to suggest that the average student writing score differs between levels of parental education.

## Part Three

#### Establish Variables and Hypotheses

Categorical Variables: Parental level of education, Race/ethnicity group

**Null hypothesis**: There is no relationship between a student's parental level of education and their race/ethnicity.

**Alternative hypothesis**: There is a relationship between a student's parental level of education and their race/ethnicity.

#### Contingency Table

```{r}
table <- table(studentdata$parental.level.of.education, studentdata$race.ethnicity)
```

#### Conduct Chi-Squared Test for Association

```{r}
chisq_result <- chisq.test(table)

chisq_result$p.value

chisq_result$residuals

corrplot(chisq_result$residuals, is.cor=F)
```

The p-value from this test is 0.07911, which is greater than the level of significance (0.05). There isn't sufficient statistical evidence to suggest that there is an association between a student's parental level of education and their race/ethnicity.
